Nearest Neighbour Methods for Length of Stay Prediction

### 2015-02-08 papers/paper.tex
Add more content to results and discussion section

\subsection{Basic Algorithm}
The $k$-nearest neighbour classifier is an example of an
\textit{instance-based} learning algorithm. This is because it does not attempt
to deduce or
generalise a relationship between the features and the class: it simply stores
all training examples and classifies an unseen example by finding the $k$
``nearest'' examples (or \textit{instances}) and assigning the majority class
to the new example. In the simplest form of the algorithm, $k$ is chosen to be
1 and the Euclidean distance is used to measure the closeness of neighbours:
\begin{equation*}
\mathrm{Distance} = \sqrt{(x_1-y_1)^2 + (x_2-y_2)^2 + \ldots + (x_n-y_n)^2}
\end{equation*}
where $x_i,y_i$ are the values of the $i$-th feature for two different feature
vectors (or instances).

@@

\subsection{Classification Algorithms}
We tested the performance of
support vector machines, multi-layer perceptrons, $k$-nearest neighbour with 1
and 20 nearest neighbours, K* and logistic regression, as well as our proposed
Ranked Distance Nearest Neighbour -- a total of 7 different classifiers.

@@

\section{Results and Discussion}
\label{sec:results}
The best accuracy was 77.81\%, achieved using the $k$-NN algorithm
with 1 nearest neighbour, a discretised data set, and feature selection using a
C4.5 wrapper. The best AUC of 0.846 was achieved using logistic regression with
features selected by a correlation coefficient threshold of 0.1, and again
using the discretised data set.

@@

It is worth noting that our best accuracy, using $k$-NN with 1
nearest neighbour, shows a statistically significant improvement upon the accuracy
that they achieved by 2.75\%
while using only 11 features (compared to the 19 they used). These 11 features
make up only 14.1\% of all features in the data set, and performed better than
the 11 features selected manually by the domain expert.

@@

The best accuracy and AUC results for
each classifier resulted from a reduced feature set obtained through a
particular feature selection method. There was at least one feature selector
for each classifier that managed to outperform the same classifier with
no feature selection. This is likely due to the large number of features in
the original data -- after cleaning and preliminary removal of redundant
features, we were still left with over 50. Many of these would still likely
be irrelevant to predicting the class.
Out of 20 combinations of classifiers
and discretisation (the 10 classifiers with and 10 without
discretisation), the C4.5 wrapper method resulted in the best
accuracy for 7 of these combinations, followed by 5 using information gain
and 4 using the correlation coefficient and very few for the other methods.
Under no combination was 1R the best feature selection method to use, as other
methods produced better accuracy.

@@

@@

>>>>

\subsection{Basic Algorithm}
The $k$-nearest neighbour ($k$-NN) classifier is an example of an
\textit{instance-based} learning algorithm. This is because it does not attempt
to deduce or
generalise a relationship between the features and the class: it simply stores
all training examples and classifies an unseen example by finding the $k$
``nearest'' examples (or \textit{instances}) and assigning the majority class
to the new example. In the simplest form of the algorithm, $k$ is chosen to be
1 and the Euclidean distance is used to measure the closeness of neighbours:
\begin{equation*}
\mathrm{Distance} = \sqrt{(x_1-y_1)^2 + (x_2-y_2)^2 + \ldots + (x_n-y_n)^2}
\end{equation*}
where $x_i,y_i$ are the values of the $i$-th feature for two different feature
vectors (or instances).

@@

\subsection{Classification Algorithms}
We tested the performance of
support vector machines (SVMs), multi-layer perceptrons (MLPs), $k$-nearest
neighbour with 1
and 20 nearest neighbours, K* and logistic regression (LR), as well as our
proposed
Ranked Distance Nearest Neighbour (RD) -- a total of 7 different classifiers.

@@

\section{Results and Discussion}
\label{sec:results}

\subsection{Overall Performance}
The best accuracy was 77.81\%, achieved using the $k$-NN algorithm
with 1 nearest neighbour, a discretised data set, and feature selection using a
C4.5 wrapper. The best AUC of 0.846 was achieved using logistic regression with
features selected by a correlation coefficient threshold of 0.1, again
using the discretised data set.

@@

Recall that in order for our work to be comparable to that of previous work on
trauma LOS prediction by Dinh et al. \cite{Dinh2013a}, we use the same data set
that they used, and also tested the features they used as one of the manual
feature selection methods.
From Table \ref{tab:overall-tr-acc} we can see that
the best accuracy achieved using their features is 75.06\% with logistic
regression.
It is worth noting that our best accuracy, using $k$-NN with 1
nearest neighbour, shows a statistically significant improvement of 2.75\% upon
the accuracy achieved by Dinh et al.
while using only 11 features (compared to the 19 they used). These 11 features
make up only 14.1\% of all features in the data set, and performed better than
the 11 features selected manually by the domain expert.

@@

\subsection{Comparison of Feature Selection Methods}
The best accuracy and AUC results for
each classifier resulted from a reduced feature set obtained through a
particular feature selection method. There was at least one feature selector
for each classifier that managed to outperform the same classifier with
no feature selection. This is likely due to the large number of features in
the original data -- after cleaning and preliminary removal of redundant
features, we were still left with over 50. Many of these would still likely
be irrelevant to predicting the class.
Out of 20 combinations of classifiers
and discretisation (the 10 classifiers with and 10 without
discretisation), the C4.5 wrapper method resulted in the best
accuracy for 7 of these combinations, followed by 5 using information gain
and 4 using the correlation coefficient and very few for the other methods.
Under no combination was 1R the best feature selection method to use, as other
methods produced better accuracy.

@@

Note that regardless of whether we consider accuracy or AUC to be more
important in evaluating how our classifiers performed, neither the baseline
feature set from Dinh et al. \cite{Dinh2013a} nor the features suggested by
the domain expert (both of which were manually selected) yielded better
results than automatic feature selection in all but two cases. We did find
that expert selection produced the best accuracy result for 1R (equal with
correlation coefficient selection, but used less features) and that the
baseline feature set produced the best accuracy for Na\"{i}ve Bayes. Neither
feature selection method produced the best AUC result for any classifier.
This is a surprising result because intuitively and in the literature it
has been stated that in a highly specialised field such as medicine, the domain
experts would be best-equipped to make judgements about which features to
use \cite{Witten2005}, and these should give reasonable results. However,
our results show that in attempting to classify patients according to their
LOS, we should consider both manual and automatic approaches.

@@

\subsection{Comparison of Classifiers}
Figures \ref{fig:tr-nothreshold} and \ref{fig:tr-threshold} compare the
accuracy and AUC of all classifiers under each feature selection method for
the trauma data set.
The methods that do not select features based upon a specified threshold
(manual methods as well as CFS, C4.5 wrapper and no feature selection)
are graphed in Figure \ref{fig:tr-nothreshold}, and the methods that select
based upon thresholds (correlation, information gain and 1R) are graphed
in Figure \ref{fig:tr-threshold}, showing varying accuracy and AUC as the
threshold is changed.

In terms of classification accuracy, we did not find a lot of variation
between all classifiers, with Zero-R producing 66.91\% accuracy and 77.81\%
being the best we achieved -- an improvement of just over 10\%. The AUC of
the 11 classifiers we tested ranged from 0.498 with Zero-R to 0.846 with
logistic regression. Logistic regression, the \textit{de facto} method
used in LOS classification, performed consistently better than the others
across all feature
selection methods, but only when the feature set was not heavily reduced by
the feature selector: in these cases, such as the results from Figure
\ref{fig:tr-nothreshold-cfs-acc} showing classifier performance with features
selected by CFS, C4.5 decision trees and nearest
neighbour methods have higher accuracy than logistic regression. This is also
apparent from the graphs in Figure \ref{fig:tr-threshold}: for all three
feature selection methods, as we increase the threshold, logistic regression
drops in performance after the initial few thresholds and is outperformed by
C4.5 and the various nearest neighbour classifiers ($k$-NN with 1 and 20
neighbours and K*), as well as our Ranked Distance algorithm.
This is true for both accuracy and AUC.

We found that in most cases, the SVM did not achieve a higher accuracy or
AUC than logistic regression with the same features even though SVMs are
considered the state-of-the-art in classification algorithms. Additionally,
as the feature set was reduced, the discriminating ability of SVMs (as
indicated by the AUC) decreased more than their accuracy: this is particularly
noticeable in Figure \ref{fig:tr-threshold-oner}, where the AUC of the SVM
drops sharply to below the AUC of even the 1R classifier at the 67.905\%
threshold, where there are
only 4 features. SVMs have not been used in a LOS classification problem like
ours, so there will need to be further work carried out in order to assess the
suitability of using SVMs to classify patient LOS.

Although the MLP has been investigated in studies attempting to identify
factors associated with longer LOS or mortality for trauma patients
\cite{Hunter2000,McGonigal1993}, leading to improved results with the neural
network, we did not see the MLP outperform all other classifiers with respect
to accuracy or AUC for any feature selection method for this data set. In fact,
the MLP often performed no better than any of the nearest neighbour
classifiers, which use a straightfoward mechanism of classifying unseen
examples. Even if the MLP had managed to achieve superior accuracy or AUC, the
``black box'' nature of its prediction decisions are a barrier for their
widespread use in medical decision-making.

Nearest neighbour approaches appear to have performed well on this data set in
both accuracy and AUC, especially in comparison to more sophisticated
classifiers and when the feature set is significantly reduced. From the graphs
in Figure \ref{fig:tr-threshold}, we can see that all the nearest neighbours
approaches perform similarly when the number of features is decreased, but our
Ranked Distance algorithm showed a statistically significant improvement over
other NN algorithms. This
improvement is not apparent at smaller feature sets, where using the standard
$k$-NN algorithm with 1 or 20 neighbours performs slightly better.
The performance of NN we observed in our results is
an interesting finding because nearest neighbours have not been commonly
applied in predicting the LOS, but there is intuitive appeal in the
case-by-case nature of how these algorithms classify new examples that should
be explored further. Unlike MLPs and SVMs, prediction decision of nearest
neighbours are more transparent and understandable. However, the major drawback
of NN algorithms is their storage requirement and the computational time
required to search for the nearest neighbours to make the classification.

There is a limitation to our findings that needs to be noted: we did not
attempt to tune the parameters of the classifiers in order to find the ones
that resulted in the best performance, so our statements are based on the
results of the default parameters of each classifier.
MLPs in particular are sensitive to
the network architecture, so it is likely that its accuracy and AUC could be
improved if the parameters were selected to optimise its performance on this
data set.
